Formal Verification Framework Based on Symbolic Execution for Smart Contract

### 2023-12-28 papers/static-checker/ecoop2024/ecoop2024.tex
write oracle part

\subsubsection{approve}
The crucial property to ensure the correctness of the approve entrypoint is that changing the allowance value from a non-zero value to another non-zero value is forbidden to prevent a corresponding attack vector \ref{}. The basic idea of this attack is that when a non-zero value update is allowed, the \lstinline/spender/ can attempt to spend the entire allowance before the new amount is updated and then spend again with the new value, resulting in a larger amount than actually allowed. This means the contract call must fail when both the parameter \lstinline/value/ is non-zero and the allowance of \lstinline/spender/ is also non-zero. We verify this property by generating the fail condition, and the result is that the non-zero value update is recorded as a fail condition, as shown in Listing \ref{}. This indicates that the implementation ensure the non-zero value update prevention.

The other property is that in case the update is allowed, then the updated value needs to be recorded correctly in the storage. This property can be divided into two subproperties, which can be checked separately. The first property is that when the parameter \lstinline/value/ is zero, then the corresponding allowance value of the \lstinline/spender/ from the owner (which is \SENDER\ in this call) should be zero. This subproperty is specified as shown in Listing \ref{}. The second property is that when \lstinline/value/ is non-zero, requiring the allowance of the \lstinline/spender/ to be zero, then the \lstinline/value/ should be recorded in the storage after the call. To check this property, we specify the property as the pre and post conditions, as illustrated in Listing \ref{}.

@@

\begin{lstlisting}[float,captionpos=b,caption={Specification for the \lstinline/getBalance/ entrypoint},label={lst:specification-get-balance},numbers=left]
Input := Pair (Pair (owner: address) (c: contract nat))
  (Pair (ledger: big_map address (pair nat (map address nat))) 
         (Pair (admin: address) 
               (Pair (paused: bool) (total_supply: nat))))
Pre-condition :=  Get_big_map (owner, ledger) = 
  Some value: option (pair nat (map address nat)) |
  value  = Pair (k: nat) (m: (map address nat))
Post-condition := post_ledger = ledger |
  post_admin = admin |
  post_paused = paused |
  post_total_supply = total_supply |
  post_Balance = Balance | 
  Transfer_token (k, Amount, c) ;

Pre-condition :=  Get_big_map (owner, ledger) = 
  None option (pair nat (map address nat)) 
Post-condition := post_ledger = ledger |
  post_admin = admin |
  post_paused = paused |
  post_total_supply = total_supply |
  post_Balance = Balance | 
  Transfer_token (0, Amount, c) ;
\end{lstlisting}
\subsubsection{Other entrypoints}
The contract provides five additional entrypoints: \lstinline/mint/ to mint you own, \lstinline/burn/ to burn existing tokens, \lstinline/setAdministrator/ and \lstinline/getAdministrator/ to set up or retrieve the administrator of the smart contract, and \lstinline/setPause/ to pause the contract. To generate fail conditions, we detect that \lstinline/transfer/ or \lstinline/approve/ calls fail if the contract is in a paused state. Users or dapps should be aware of this when interacting with the smart contract.

@@

All of the above-mentioned properties are specified and verified in our model, and the results show that these properties are guaranteed. This confidence assures us of the correctness of the smart contract code.
\subsection{Kolibri Oracle Contract}
\section {Conclusion}
\section {Appendix}
\end{document}

>>>>

\subsubsection{approve}
The crucial property to ensure the correctness of the approve entrypoint is that changing the allowance value from a non-zero value to another non-zero value is forbidden to prevent a corresponding attack vector \ref{}. The basic idea of this attack is that when a non-zero value update is allowed, the \lstinline/spender/ can attempt to spend the entire allowance before the new amount is updated and then spend again with the new value, resulting in a larger amount than actually allowed. This means the contract call must fail when both the parameter \lstinline/value/ is non-zero and the allowance of \lstinline/spender/ is also non-zero. We verify this property by generating the fail condition, and the result is that the non-zero value update is recorded as a fail condition, as shown in Listing \ref{lst:approve-udstz-contract}. This indicates that the implementation ensure the non-zero value update prevention.
\begin{lstlisting}[float,captionpos=b,caption={Fail conditions for the \lstinline/approve/ entrypoint},label={lst:approve-udstz-contract},numbers=left]
1. paused: 'TokenOperationsArePaused'
2. not paused &&
   get_big_map (Sender, ledger) = Some sbvar &&
   sbvar = Pair sbvar1 sbvar2 &&
   get_map (spender, sbvar2) = Some sbvar3 &&
   not (value = 0) &&
   not (sbvar3 = 0): 'UnsafeAllowanceChange'
\end{lstlisting}
The other property is that in case the update is allowed, then the updated value needs to be recorded correctly in the storage. This property can be divided into two subproperties, which can be checked separately. The first property is that when the parameter \lstinline/value/ is zero, then the corresponding allowance value of the \lstinline/spender/ from the owner (which is \SENDER\ in this call) should be zero. This subproperty is specified as shown in Listing \ref{lst:specification-approve}. The second property is that when \lstinline/value/ is non-zero, requiring the allowance of the \lstinline/spender/ to be zero, then the \lstinline/value/ should be recorded in the storage after the call. To check this property, we specify the property as the pre and post conditions, as illustrated in Listing \ref{}.

\begin{lstlisting}[float,captionpos=b,caption={Specification for the \lstinline/approve/ entrypoint},label={lst:specification-approve},numbers=left]
Input := Pair (Pair (owner: address) (c: contract nat))
  (Pair (ledger: big_map address (pair nat (map address nat))) 
         (Pair (admin: address) 
               (Pair (paused: bool) (total_supply: nat))))
Pre-condition := not paused |
   get_big_map (Sender, ledger) = 
   Some (v: option (pair nat (map address nat))) | 
   v = Pair (n1: nat) (m: map address nat) |
   get_map (spender, m) = Some (n2: nat) |
   not (n2 = 0) |
   (value = 0)
Post-condition := get_big_map (Sender, post_ledger) = 
   Some (post_v: option (pair nat (map address nat))) | 
   post_v = Pair (post_n1: nat) (post_m: map address nat) |
   get_map (spender, post_m) = Some (post_n2: nat) |
   (post_n2 = 0) ;

Pre-condition := not paused |
   get_big_map (Sender, ledger) = 
   Some (v: option (pair nat (map address nat))) | 
   v = Pair (n1: nat) (m: map address nat) |
   get_map (spender, m) = Some (n2: nat) |
   (n2 = 0) |
   not (value = 0)
Post-condition := get_big_map (Sender, post_ledger) = 
   Some (post_v: option (pair nat (map address nat))) | 
   post_v = Pair (post_n1: nat) (post_m: map address nat) |
   get_map (spender, post_m) = Some (post_n2: nat) |
   (post_n2 = value) ;
\end{lstlisting}

@@

\begin{lstlisting}[float,captionpos=b,caption={Specification for the \lstinline/getBalance/ entrypoint},label={lst:specification-get-balance},numbers=left]
Input := Pair (Pair (owner: address) (c: contract nat))
  (Pair (ledger: big_map address (pair nat (map address nat))) 
         (Pair (admin: address) 
               (Pair (paused: bool) (total_supply: nat))))
Pre-condition :=  get_big_map (owner, ledger) = 
  Some v: option (pair nat (map address nat)) |
  v  = Pair (k: nat) (m: (map address nat))
Post-condition := post_ledger = ledger |
  post_admin = admin |
  post_paused = paused |
  post_total_supply = total_supply |
  post_Balance = Balance | 
  transfer_token (k, Amount, c) ;

Pre-condition :=  get_big_map (owner, ledger) = 
  None option (pair nat (map address nat)) 
Post-condition := post_ledger = ledger |
  post_admin = admin |
  post_paused = paused |
  post_total_supply = total_supply |
  post_Balance = Balance | 
  transfer_token (0, Amount, c) ;
\end{lstlisting}
\subsubsection{Other entrypoints}
The contract provides five additional entrypoints: \lstinline/mint/ to mint you own, \lstinline/burn/ to burn existing tokens, \lstinline/setAdministrator/ and \lstinline/getAdministrator/ to set up or retrieve the administrator of the smart contract, and \lstinline/setPause/ to pause the contract. To generate fail conditions, we detect that \lstinline/transfer/ or \lstinline/approve/ calls fail if the contract is in a paused state. Users or dapps should be aware of this when interacting with the smart contract.

@@

All of the above-mentioned properties are specified and verified in our model, and the results show that these properties are guaranteed. This confidence assures us of the correctness of the smart contract code.
\subsection{Kolibri Oracle Contract}
The potential of smart contracts is constrained by the inability to access external data sources. To address this limitation, oracles serve as a crucial link between external sources and the blockchain network.  Blockchain oracles plays a pivotal role in furnishing external data for smart contracts. Certain essential properties must be satisfied by an oracle contract, including data approvability. This case study aims to illustrate the process of specifying and verifying an oracle contract by the static checker, using the Kolibri oracle contract as an exemple.

The Kolibri oracle contract,  residing at the address KT1Tj6xknbwjyK5gyfESNR6WESBcP3yX1mmj, ensures the accuracy of external data for the Tezos-based stablecoin, Kolibri, by leveraging the Harbinger Price Feed. The contract initiates calls to a Harbinger Normalizer contract, check  on the returned values, normalizes them to suit the Kolibri system, and subsequently relays them to the client. The smart contract's parameters and storage details are outlined in Listing \ref{lst:kolibri-oracle-contract}. Key attributes stored by the Oracle include:
\begin{itemize}
\item \lstinline/harbingerContractAddress/: the address of the Harbinger Price Feed contract.
 \item    \lstinline/state/: a natural number indicating the current state of the contract. The oracle contract operates as a simple state machine with two states: (1) \lstinline/IDLE/ and (2) \lstinline/WAITING_FOR_HARBINGER/.
 \item    \lstinline/clientCallback/: a contract featuring a \TNAT\ entrypoint that stores state when the oracle awaits Harbinger data.
 \item    \lstinline/governorContractAddress/: the address of the governor contract.
 \item    \lstinline/maxDataDelaySec/: the maximum acceptable age of data received from Harbinger, measured in seconds.
\end{itemize}
The oracle offers various entrypoints to interact with the smart contract, including:
\begin{itemize}
\item    \lstinline/getXtzUsdPrice/: facilitates the retrieval, sanity check, and normalization of Harbinger data.
\item    \lstinline/getXtzUsdPrice_callback/: serves as a callback mechanism to receive data from Harbinger.
\item    \lstinline/default/: always intentionally fails to prevent the transfer of XTZ into the contract.
\item    \lstinline/setMaxDataDelaySec/: enables the setting of the maximum acceptable age for data received from Harbinger.
\item    \lstinline/setGovernorContract/: allows the setting of the Governor's address.
\end{itemize}

\begin{lstlisting}[float,captionpos=b,caption={Kolibri oracle contract},label={lst:kolibri-oracle-contract},numbers=left]
parameter
  ( or
    ( or
      ( unit %default )
      ( contract %getXtzUsdRate nat ) )
    ( or
      ( pair %getXtzUsdRate_callback string 
      ( pair timestamp nat ) )
      ( or
        ( address %setGovernorContract )
        ( nat %setMaxDataDelaySec ) ) ) ) ;
storage
  ( pair
    ( pair ( option %clientCallback address ) 
          ( address %governorContractAddress ) )
    ( pair ( address %harbingerContractAddress ) 
          ( pair ( nat %maxDataDelaySec ) ( int %state ) ) ) ) ;
\end{lstlisting}

\section {Conclusion}
\section {Appendix}
\end{document}
